---
title: "7-16-24 Class"
author: "Courtney Hodge"
date: "2024-07-16"
output: html_document
---
Call file.choose() to find the directory of a file on your machine
```{r}
Classdata <- read.csv("C:\\Users\\hodge\\Downloads\\ClassData.csv")
```

```{r}
library(tidyverse)
```

```{r}
OurData5 <-data |> mutate(sleep = as.numeric(Sleep_Hrs))

```

```{r}
#defaults to 95% CI
t.test(OurData5$sleep)
```

# Bootstrapping
```{r}
mean_calc <- function(x){
  
  return(mean(x, na.rm = T))
}

#replicate is done 10,000 times
bootstrap_means <- replicate(10000, {
  bootstraped_data <- sample(OurData5$sleep, replace = T)
  mean_calc(bootstraped_data)
})

bootstrap_df <- data.frame(bootstrap_means)

#normal dist of boostrap df
ggplot(bootstrap_df, aes(x = bootstrap_means)) +
  geom_density()
```
```{r}
quantile(bootstrap_means, c(0.025, 0.975))
#95% confidence interval is (6.969,7.298) <- note that this will change with each run
```

# Testing Hypothesis about Proportions

```{r}
prop.test(40, 100, p = 0.85, alternative = "l")
#our p value is very close to zero (2.2e-16). Since our p-value < 0.01 = alpha, you will reject the null hypothesis!

#note, whenever your p-val < H0, you say that it is a statistically significant result
```
A final project group member tells you that all UVa students sleep for 6
hours on average. Use students from this class as a random sample. Do
you believe your classmate - test your conjecture that all UVa students
sleep more than 6 hours on average? Test at α = 0.05.

```{r}
#default is 2 sided 
t.test(OurData5$sleep, mu = 6, alternative = "g")
```
> We will reject the null because p val < alpha.


```{r}
#list the (successes, the sample size)
#Slide pack 2, pg 8
prop.test(c(7, 15), c(15, 19))
#we are 95% confident that the proportion of customers who would select A over B is (-0.694, 0.04)

# we can therefore not make a conclusion about this
```
```{r}
nba_data <- read.csv("C:\\Users\\hodge\\Downloads\\nba.csv")
```

```{r}
home_count <- count(nba_data$home_away)
```

