---
title: "7-25-24 Class"
author: "Courtney Hodge"
date: "2024-07-25"
output: html_document
---

# Inference on the linear model
```{r}
library(tidyverse)
```


```{r}
Startups <- read.csv("C:\\Users\\hodge\\OneDrive - Baylor University\\Desktop\\UVA Coding Folder\\STAT 6021\\Startups.csv")
```

```{r}
model1 <- lm(Profit~R.D.Spend + Administration + Marketing.Spend, data = Startups)

summary(model1)
```
We now want to investigate the p-value

R.D.Spend is less than 0.05, so we are good with this value b/c the p-value is 2e-16

Since Administration's p-value is 0.602, we can't use it to predict the model, so we don't care about it.

With Marketing.Spend, it is our call to make just because we don't have that many predictors. It is technically greater than 
($\alpha = 0.05$).So, we are keeping it.

Normally, what we would do is recalculate the model and get different p-values, which might end up changing the model. Let's try that building a second model.

```{r}
model2 <- lm(Profit~R.D.Spend + Marketing.Spend, data = Startups)
summary(model2)
```

Welp, we are actually getting rid of model 2. Notice that the p-values are changing. This is because the standard errors are decreasing in the summary of the output.

There's a way to automate all of the Bonferroni process. It's Called **STEPWISE REGRESSION**. There's forward selection and backwardselimination. At the beginning of the model, it will start with 1 predictor and apply the Akaike Information Criteria (AIC), which gives a metric that determines if that variable is adding too much noise into the model. If that's the case, it will drop the variable and add another.

```{r}
library(MASS)
```


```{r}
model3 <- lm(Profit~R.D.Spend + Marketing.Spend, data = Startups, method = "stepAIC")

#model 1 has all three predictors
aic<- stepAIC(model1, direction = "both")
```
We want the model with the smallest AIC

# Partial Effect Plots
added variable plots (avPlot)
```{r}
avPlots(model1)
```
> The partial effect plots shows the slope of the two residuals plot IF R.D.Spend and Asministration are both held fixed.

> The negative slope plot says every time Administration slope goes up 1 dollar, the profit decreases. 

> The last graph says that every time the marketing.spend plot goes up, the profit also increases. 

> The point of this plot is to show us the effect of each predictor on the response variable if other predictors are held fixed.
