---
title: "Class Activity 8"
author: "Courtney Hodge"
date: "2024-07-25"
output: html_document
---

# Multicollinearity

```{r}
library(tidyverse)
```

```{r}
data <- read.csv("C:\\Users\\hodge\\Downloads\\credit_data.csv")

data
```


## 1
```{r}
data <- data.frame(data)
data2 <- data[2:7]
cor_matrix <-round(cor(data2),2)
cor_matrix
```

```{r}
library(ggcorrplot)
ggcorrplot(cor_matrix, lab = T, type = "lower")
```

> There are issues with multicollinearity between Limit and Rating because their because correlation coefficients for pairwise comparisons between predictors ashould ideally be below 0.80

## 2

```{r}
library(car)
```

```{r}
model1 <- lm(Balance~Income + Limit + Rating + Cards + Age + Education, data = data)

vif(model1)
```
> After further investigation, Limit and Rating still have multicollinearity issues based on their VIF values.

```{r}
model2 <- lm(Balance~Income + Rating + Cards + Age + Education, data = data)
```

```{r}
vif(model2)
```
> After eliminating Limit, the vif values have decreased for all predictors and there are no further multicollinearity.

## 3
```{r}
model3 <- lm(Balance~Income + Rating + Cards + Age + Education, data = data)

coef(model3)
```

```{r}
new_dat <- data.frame(Income = c(-7.5431353), Rating = c(3.9366411), Cards = c(2.5641756), Age = c(-0.9018958), Education = c(2.4071949))

predict(model3, newdata = new_dat)
```

```{r}
predict(model3, newdata = new_dat, interval = "prediction", level = 0.95)
```

> We are 95% confident that a specific individual would have a predicted balance values fall between -771.5115 and -109.1669


```{r}
predict(model3, newdata = new_dat, interval = "confidence", level = 0.95)

```

> We are 95% confident that the mean predicted balance values would fall between -536.9088 and -353.7696
