---
title: "Class Activity 9"
author: "Courtney Hodge"
date: "2024-07-31"
output: html_document
---
```{r}
library(tidyverse)
```


```{r}
startups <- read.csv("C:\\Users\\hodge\\Desktop\\UVA_Coding_Folder\\Statistics-6021\\Startups.csv")
```

# 1

```{r}
model1 <- lm(Profit~ State, data = startups)
summary(model1)
```

> The model is: 

$\hat{Profit} = 103905 + (14869 * StateFlorida) + (9851 * StateNew York)$

# 2
```{r}
model2 <- lm(Profit~State + R.D.Spend, data = startups)
summary(model2)
```
# 2a
* Explore the three variables in a visualization. Also, superimpose a linear regression
line predicting Profit based on R.D.Spend. for each State.
```{r}
ggplot(startups, aes(x = Profit, y = R.D.Spend, color = State)) + geom_jitter() + geom_smooth(method = "lm", model.extract(model2), se = F)
```

# 2b
* write down the model

> The model is: 

$\hat{Profit} =  4.875e+04 + (1.164e+03 * StateFlorida) + (9.597e+00 * StateNew York) + (8.530e-01*R.D.Spend) $

# 2c

The coefficient of StateFlorida in this context is   1.164e+03 more than the California Baseline of 4.875e+04 when R.D.Spend is included in the model. 

# 3
```{r}
model3 <- lm(Profit~State * R.D.Spend, data = startups) 
summary(model3)
```

# 3a

$\hat{Profit} =  4.388e+04 + (9.242e+03 * StateFlorida x -1.151e-01) + (7.921e+03 * StateNew York x -1.153e-01)$

# 3b 

When the company is in Florida, the effect of R.D. spend on Profit is reduced 1.151e-01 times on average, and when the company is in New York, the effect of R.D.Spend on Profit is reduced 1.153e-01 times on average.

# 4

```{r}
startups$State2<-factor(startups$State, levels = c("New York", "California", "Florida"))

mod4 <- lm(Profit~State2, data = startups)
summary(mod4)
```

```{r}
startups$State3<-factor(startups$State, levels = c("Florida", "California", "New York"))

mod5 <- lm(Profit~State3, data = startups)
summary(mod5)
```
State is not a useful predictor based on the p-values. When we change the reference category for state, it still remains not significant, therefore, the predictor category for State is not useful.