---
title: "Class Activity 2"
author: "Courtney Hodge"
date: "2024-07-17"
output: html_document
---
```{r}
library(tidyverse)
```

```{r}
AB_Test_data <- read.csv("C:\\Users\\hodge\\Downloads\\WA_Marketing-Campaign.csv")
```

```{r}
Promotion3_data<-subset(AB_Test_data, Promotion==3)
```

#1.a

```{r}
mean_calc <- function(x){
  
  return(mean(x, na.rm = T))
}

#replicate is done 10,000 times
bootstrap_means <- replicate(1000, {
  bootstraped_data <- sample(Promotion3_data$SalesInThousands, replace = T)
  mean_calc(bootstraped_data)
})

bootstrap_df <- data.frame(bootstrap_means)


quantile(bootstrap_means, c(0.025, 0.975))
```

> The mean interval falls between 52.98 and 57.75

#1.b 

```{r}
t.test(Promotion3_data$SalesInThousands)
```
> We are 95% Confident that the mean sale amount for all restuarants that adopt Promotion 3 falls between 52.95221 and 57.77673 measured in the thousands

#1.c

> based on the similar intervals of part a and b, the widths of the Bootstrap and the t-intervals are about the same. Their measurements for mean sale amounts both fall a couple tenths of points above 52 and below 58.

#2
```{r}
Promotion_2_3<-subset(AB_Test_data, Promotion==2 | Promotion==3)
```
 
```{r}
t.test(SalesInThousands~Promotion, data = Promotion_2_3)
```

> We advise them to use Promotion 3 over Promotion 2 and, on average, we notice that with Promotion 3, companies can expect to generate between 4.8 and 11.3 thousand more sales over Promotion 2. 