---
title: "Class Activity 4"
author: "Courtney Hodge"
date: "2024-07-19"
output: html_document
---

```{r}
library(tidyverse)
```


## Startups Data
```{r}
BookData <- read.csv("C:\\Users\\hodge\\Downloads\\Books_Data_Clean.csv")

```

```{r}
ggplot(BookData, aes(x = Author_Rating, y = gross.sales, fill = Author_Rating )) +
  geom_boxplot()
```
> zooming into graph

```{r}
ggplot(BookData, aes(x = Author_Rating, y = gross.sales, fill = Author_Rating )) +
  geom_boxplot() + ylim(0, 5000)
```


> Hypotheses

H0: Mu A = Mu B = Mu C = Mu D
HA: At least one of them are different

```{r}
anova <- aov(gross.sales~Author_Rating, data = BookData)

summary(anova)
```
> Since our p-value of 3.03e-11 is incredibly small and less than 5%, then p-val < 0.05 means that we reject the null-hypothesis.

> multiple hypothes testing

MuFamous-MuExcellent

MuIntermediate-MuExcellent 

MuNovice-MuExcellent

MuIntermediate-MuFamous

MuNovice-MuFamous

MuNovice-MuIntermediate 

```{r}
TukeyHSD(anova, conf.level = 0.95)
```

MuFamous-MuExcellent  --> - > + (inconclusive)

MuIntermediate-MuExcellent -->  + > + 

MuNovice-MuExcellent --> + > + 

MuIntermediate-MuFamous --> - > + (Incolcusive)

MuNovice-MuFamous --> + > +

MuNovice-MuIntermediate --> - > + (incolcusive)

```{r}
plot(TukeyHSD(anova, conf.level = 0.95))
```

> Based on the analysis above, it looks like drug A is the most effective drug in this experiment. we found that by looking at the lower and upper bounds and noticing that A is positive, but C and B are negative.

> We have enough statistical evidence to conclude that the mean pain rating for Drug A is much lower than Drugs B and C. Becuase this is an experiment, Drug A is more effective in treating migraine headaches compared to Drugs B and C